---
title: "Ch-2"
author: "David Crandall"
date: "May 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## College Data 

The following are exercises from Ch.2 of *An Introduction to Statistical Learning*

[Description of Cols](https://vincentarelbundock.github.io/Rdatasets/doc/ISLR/College.html)

> Statistics for a large number of US Colleges from the 1995 issue of US News and World Report.

```{r}
college = read.csv("College.csv")
summary(college)
```

## Create row names

As you can see from the summary output, we have a column `X` with our school names. It would be nice to have these be our row names, as opposed to data. For this we can use the `rownames` function, which sets the row names on our `college` dframe using the values from the first column, which has the names of our schools.

```{r}
rownames(college) = college[,1]
summary(college)
```

If you call `fix(college)` you will see a new col labelled `row.names` but we still have that pesky `X` col, so let's get rid of that.

```{r}
college = college[,-1]
summary(college)
```

## Create a new col of *Elite* Schools

First, create a vector of length equal to the number of rows in our college dframe For this, we can use `rep`, which takes a value to replicate and the number of times to replicate this value. `nrow` can be used here to return the number of rows in our matrix.

```{r}
Elite = rep("No", nrow(college))
```

Now we need to assign "Yes" to the Elite schools. Elite schools are defined as schools where over 50% of new students are from the Top 10% of their HS class. We also need to tell R that this is categorical data, which we can do using `as.factor(Elite)`. Then finally, we will create new `college` dframe with our `Elite` vector added.

```{r}
Elite[college$Top10perc > 50] = "Yes"
Elite = as.factor(Elite)
college = data.frame(college, Elite)
summary(college)
```

Very Cool, now let's so some visualization

## Data Visualization

### Plot
Let's see how many PhD's work at the elite schools relative to non-elite schools. For this we can use `plot()`

```{r}
plot(college$Elite, college$PhD, xlab="Elite", ylab="% Faculty with a PhD")
```

Well, I suppose that's not much of a surprise -- faculty at elite schools are more like to have a PhD.

I wonder if books are more expensive at the elite schools and what the student/faculty ratios are like. Let's put these plots on the same figure too, using `par()`

```{r}
par(mfrow=c(1,2))

plot(college$Elite, college$Books, xlab="Elite", ylab="Est. Books Cost")

plot(college$Elite, college$S.F.Ratio, xlab="Elite", ylab="student/faculty ratio")
```

Eh, marginal differences. There has to be something a bit juicier.

```{r}
par(mfrow=c(1,2))

plot(college$Elite, college$perc.alumni, xlab="Elite", ylab="% alums who give $")

plot(college$Elite, college$Expend, xlab="Elite", ylab="$ spent on students")
```



```{r}
plot(college$Elite, college$Grad.Rate, xlab="Elite", ylab="$ spent on students")
```

